### 2011-07-13 papers/paper1/design.tex

\section{Design and Implementation\label{design}}

\begin{itemize}
  \item Problem
  \item Design forces
  \item Solution
  \item Consequences
\end{itemize}