### 2011-07-22 papers/paper1/paper.tex

\documentclass[letterpaper]{article}

\usepackage{outline}

\newcommand{\comment}[1]{[#1]}

\begin{document}

\title{Implementing Distributed Systems Directly with I/O Automata in Ioa++}
\author{Justin R. Wilson \and Christopher D. Gill}
\date{}

\maketitle

\input{abstract}

\input{introduction}

\input{system_model}

\input{programming_model}

\input{design}

\input{evaluation}

\input{related_work}

\input{conclusion}

\end{document}