### 2011-07-22 papers/paper1/related_work.tex

\section{Related Work\label{related_work}}

pthreads (?)
Edward Lee - The Problem with Threads
Herb Sutter - The Free Lunch is Over
Early work on concurrency

Lee considers removing non-determinism.  We consider no shared state.