### 2011-09-03 papers/paper1/paper.tex
First part of Representation and Programming Model.

\documentclass[letterpaper]{article}

\usepackage{outline}
\usepackage{graphicx}
\usepackage{listings}

\lstset{numbers=left, basicstyle=\footnotesize}

\begin{document}

\title{Implementing Distributed Systems Directly with I/O Automata in Ioa++}
\author{Justin R. Wilson \and Christopher D. Gill}
\date{}

\maketitle

\input{abstract}

\input{introduction}

\input{component_model}

\input{design}

\input{programming_model}

\input{evaluation}

\input{related_work}

\input{conclusion}

\bibliography{bibliography}{}
\bibliographystyle{plain}

\appendix

\section{AsyncLCR automaton\label{asynch_lcr}}

\lstinputlisting[language=C++]{asynch_lcr_automaton.hpp}

\section{Channel automaton\label{channel}}

\lstinputlisting[language=C++]{channel_automaton.hpp}

\section{Unidirectional Ring Network\label{ring}}

\lstinputlisting[language=C++]{unidirectional_ring_leader_election.hpp}

\section{AsyncLCR driver\label{driver}}

\lstinputlisting[language=C++]{asynch_lcr.cpp}

\end{document}