### 2011-09-16 papers/paper1/case_study.tex
Case study.

\section{Case Study:  The TCP LCR Automaton\label{case_study}}

In this section, we explore (1) if ioa++ offers a convenient representation for I/O automata and (2) and if ioa++ contains the features necessary for building real distributed systems.
Our approach is to (1) express an automaton in ioa++ using an existing listing, (2) simulate a distributed system built using the automaton as debugging step, and (3) build a real distributed system by composing the automaton with automata for network services.
For continuity, we use the asynchronous LCR automaton developed in section~\ref{representation}.

\paragraph{Translating the asynchronous LCR automaton into ioa++.}
The first version of the code for the asynchronous LCR automaton was translated directly from a listing in~\cite{lynch1996distributed}.
The final version contains an additional action to reinitialize the automaton and modifications that allow an automaton that elected itself leader to nullify its election.
The motivation for these modifications will be explained when we discuss its use in a real system.
The listing in~\cite{lynch1996distributed} consists of approximately 24 lines in a formal language that must be translated to a production language.
The raw listing for the final version consists of 121 lines (5x).
Scheduling accounts for 22 of the 121 lines (20\%) and is the main source of code that cannot be traced back to the formal listing.
On average, every local action contributed 6 lines of scheduling code and every input action contributed 3 lines of scheduling code.

We observe that having a representation of I/O automata in a production language that closely matches a formal representation is of great benefit because it allows one to reason about an automaton directly from the source code.
For example, an invariant in the asynchronous LCR automaton is that all of the UIDs in the send queue of automaton $i$ must be greater than or equal to the UID of automaton $i$.
This invariant can be quickly checked against the constructor and four actions of the \verb+asynch_lcr_automaton+.
Upon construction, the send queue contains the UID of the automaton so the invariant holds initially.
The leader and init actions do not modify the send queue so the invariant holds trivially.
The receive action only adds a UID to the send queue if the UID is greater than the automaton's UID.
The send action removes a UID from the send queue so if the invariant is true before, it will be true after.
Reasoning about an automaton directly from the source code allows one to use formal methods to develop and debug systems.

\paragraph{Unit testing with simulation.}
After the asynchronous LCR automaton was implemented, we developed a simulation to test it.
The \verb+examples/asynch_lcr.cpp+ file contains the driver for the simulation.
The simulation creates a unidirectional ring of a specified size consisting of alternating \verb+asynch_lcr_automaton+ and \verb+channel_automaton+.
(The \verb+channel_automaton+ was also a direct translation from~\cite{lynch1996distributed}.)

The simulation showed that the \verb+asynch_lcr_automaton+ was not correct.
The first bug we encountered was a silent failure where the ring always failed to elect a leader.
The problem was that the leader action was never being scheduled so the leader action was added to the scheduling function.
The second bug manifested itself by electing the wrong leader, i.e., one whose UID was not the maximum.
A closer inspection showed that certain UIDs were being sent but never received.
The key insight into solving the problem was to understand that the automata were executing as the ring was being constructed, i.e., we failed to consider dynamic composition.
The solution was to add binding predicates to the outputs of the \verb+asynch_lcr_automaton+ and \verb+channel_automaton+.

\paragraph{From abstraction to real system.}
The next phase was to use the asynchronous LCR automaton to perform leader election in a real unidirectional ring of processes connected by TCP sockets.
The leader of the ring periodically sends a token which is forwarded by all nodes in the ring.
If a node has not received a token for a specified duration of time, it calls for an election.
This is the motivation behind the action that resets the \verb+asynch_lcr_automaton+.
Nodes listen for connections from their predecessor and attempt to connect to their successor.
Nodes can join and leave the ring at any time.
The protocol should eventually succeed for complete rings and restart when a ring becomes incomplete.
These dynamics admit situations where a ring might lose its leader or contain multiple leaders.
A ring containing multiple leaders should stabilize to having a single leader.
This is the motivation behind the \verb+asynch_lcr_automaton+ being able to withdraw its election.
A rare but possible and valid situation is a ring circulating a token without a leader, i.e., the leader sends a token, leaves the ring, and is replaced before the token makes a full cycle.

The \verb+tcp_lcr_automaton+ which is listed in \verb+examples/tcp_lcr.cpp+ (484 lines) of the ioa++ package implements the protocol just described.
The \verb+tcp_lcr_automaton+ is a composition of a number of automata including an \verb+asynch_lcr_automaton+, an \verb+ioa::tcp_acceptor_automaton+, an \verb+ioa::tcp_connector_automaton+, and two \verb+ioa::tcp_connection_automaton+ representing the channel from the predecessor and channel to the successor.
The TCP automata provided by the ioa++ package are ordinary automata implemented using file descriptors and the techniques described in section~\ref{design}.
The \verb+tcp_lcr_automaton+ uses the techniques for managing dynamic constellations described in section~\ref{representation}.

The \verb+tcp_lcr_automaton+ is validation that I/O automata are a good foundation for building reusable concurrent modules.
The key idea is that all of these modules have well-defined interactions when composed while being inherently concurrent.
For example, one could substitute the \verb+asynch_lcr_automaton+ with a different automaton to realize a different leader election protocol or a different protocol altogether.
Alternatively, the \verb+tcp_lcr_automaton+ could be refactored to use a different communication mechanism, e.g., UDP.
One need only focus on the protocol implemented by an automaton when composing it with an existing system as opposed to worrying about interacting threads and race conditions.

%% \begin{figure}
%% \center
%% \includegraphics[width=\textwidth]{tcp_lcr_automaton}
%% \caption{TCP LCR automaton.}
%% \label{tcp_lcr_automaton}
%% \end{figure}

%% \paragraph{Dynamics in the TCP LCR automaton.}
%% A \verb+tcp_lcr_automaton+ continuously attempts to complete the ring by accepting a connection from its predecessor or connecting to its successor.
%% Accepting can fail if the socket used for listening is currently unavailable.
%% Connecting can fail if the successor automaton does not exist and is therefore not listening for incoming connections.
%% The connections themselves will fail when nodes leave the ring.
%% The \verb+tcp_lcr_automaton+ also contains constraints on the order in which automata are created with respect to each other and with respect to which actions are executed.
%% For example, the predecessor connection must be created before it is passed to the acceptor via the accept action.
%% Similarly, the successor connection must be created before it is passed via a constructor to the connector.

%% Conclude with our recommendations for programming, highlight need for exploration.

%% Tips for Writing Programs with I/O Automata
%% 1. Make sure that a parameter exists.
%% 2. Move logic in observe to local actions.
%% 3. Avoid observing an arbitrary number of managers.
%% 4. Do not bind to managers allocated on the stack.
%% 5. Stop when an error is encoutered and report it.
%% 6. Make sure output actions are bound.
%% 7. Make sure local actions are scheduled.
%% 8. Do not pass pointers---use const_shared_ptr.
%% 9. Check preconditions.
%% 10. Check effects.
%% 11. Order clauses of preconditions for efficiency.
%% 12. Automata that should be destroyed at the same time should be created at the same time and vice versa.  (Group using a parent.)

%% \begin{outline}
%% \item State
%%   \begin{outline}
%%   \item There is no shared state in a 
%%   \item Local state only
%%   \item Shared state
%%     \begin{outline}
%%       \item Impossible in distributed systems
%%       \item Dangerous in local systems
%%     \end{outline}
%%   \end{outline}
%% \item Communication
%%   \begin{outline}
%%   \item Atomic asynchronous message passing
%%   \item Network sets size of atom (UDP)
%%   \item Can build reliable streams (TCP)
%%   \item Local equivalent is passing a value
%%   \item Model should lend itself to writing protocols
%%   \end{outline}
%% \item Asynchrony
%%   \begin{outline}
%%     \item Model must have natural support for asynchrony, i.e., event-based
%%     \item Leads to a more efficient implementation because changed state and enabled actions become obvious
%%   \end{outline}
%% \item Concurrency
%%   \begin{outline}
%%     \item Reason about systems using non-deterministic interleaving of atomic actions
%%     \item Model should admit implementations that execute concurrently
%%   \end{outline}
%% \item Dynamics
%%   \begin{outline}
%%     \item Configuration - Edges in graph of communicating components can change at run-time.
%%       \begin{outline}
%%       \item Already required in distributed settings
%%       \item Not addressed in formal models
%%       \end{outline}
%%     \item Extension - Nodes in graph of communicating components can change at run-time.
%%   \end{outline}
%%   \item Reflection
%% \end{outline}

%% I/O Automata
%% \begin{itemize}
%%   \item Compare with UNITY
%%   \item Compare with esterel
%%   \item Compare with pi calculus
%%   \item Compare with Ptolemy
%% \end{itemize}