### 2011-09-22 papers/paper1/paper.tex

\documentclass[10pt, conference, compsocconf]{IEEEtran}

\usepackage{cite}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{url}
\usepackage{amsmath}

\lstset{language=C++, basicstyle=\footnotesize}

\newif\ifjournal
\journalfalse

\begin{document}

\title{Implementing Distributed Systems Directly with I/O Automata in Ioa++}
\author{
  \IEEEauthorblockN{Justin R. Wilson and Christopher D. Gill}
  \IEEEauthorblockA{Department of Computer Science and Engineering\\
    Campus Box 1045, Washington University\\
    One Brookings Drive\\
    St. Louis, MO 63130\\
    {wilsonj,cdgill}@cse.wustl.edu}
  \and
  \IEEEauthorblockN{Guia-Catalin Roman}
  \IEEEauthorblockA{TODO:  Dr. Roman's address}
}

\maketitle

\input{abstract}

\input{introduction}

\input{component_model}

\input{design}

\input{programming_model}

\input{case_study}

\input{evaluation}

\input{related_work}

\input{conclusion}

\bibliographystyle{IEEEtran}
\bibliography{bibliography}{}

\end{document}