### 2011-09-23 papers/paper1/related_work.tex
Related work.

\section{Related Work\label{related_work}}

This work grew out of the frustration encountered when trying to develop distributed systems with threads.
Lee elaborates on the problems faced when using threads for concurrency in~\cite{lee2006problem}.
These frustrations will increase as concurrent programming is the only foreseeable way to increase performance given the leveling trend of clock frequencies and the advent of multi-core processors\cite{sutter2005software}.

The idea to implement a formal model to simplify concurrent programming was inspired by the application of structured programming~\cite{dijkstra1968letters} to programming language design.
The I/O automata model of Lynch and Tuttle\cite{lynch1987hierarchical} was selected because it supports features we believe to be fundamental to developing distributed systems.
\cite{lynch1996distributed} shows how I/O automata can be applied to a number of problems in distributed computing.
Event-based proof techniques~\cite{manna1992temporal},~\cite{lamport1978time} can be used to reason about I/O automata.

Comparing I/O automata to every model for asynchronous, concurrent, and distributed computing is beyond the scope of this paper so we focus on exemplars.
The I/O automata model is based on state transitions which matches the imperative style that dominates modern concurrent programming.
This is similar to the motivation for UNITY~\cite{chandy1988parallel} and is in contrast to functional models such as the $\pi$-calculus~\cite{milner1992calculus} and actors~\cite{agha1986actors}.
The I/O automata model assumes independent state with explicit communication like communicating sequential processes (CSP)~\cite{hoare1978communicating} but unlike threads~\cite{lee2006problem} and the UNITY model~\cite{chandy1988parallel}.
All models for asynchronous, concurrent, and distributed computing admit non-determinism.
Lee points out that the difficulties of thread-based programming come from a need to ``prune'' non-determinism when paired with shared state and adopts the view that software components should be deterministic with respect to concurrency save a few that introduce non-determinism in a controlled way~\cite{lee2006problem}.
Programs in UNITY are composed by concatenating program texts and resolving shared variables~\cite{chandy1988parallel}.
For this reason, they share many of the same problems as threads under composition.
I/O automata can be viewed as taking the alternate approach by admitting non-determinism while prohibiting shared state.
In I/O automata, the sequential flow control of threads~\cite{lee2006problem} and CSP~\cite{hoare1978communicating} is replaced by the non-deterministic execution of conditional atomic actions which resembles execution in the UNITY model~\cite{chandy1988parallel}.
The ability to create new automata can be compared to actor creation in the actor model~\cite{agha1986actors}.
Actors communicate asynchronously by name using a buffered mail system while I/O automata communicate synchronously and anonymously according to composition.

Ioa++ can be compared directly to the the IOA language~\cite{garland2003ioa} developed by Lynch et al.
IOA is programming language for I/O automata that can be used with formal verification tools.
Recently, a compiler for IOA has been developed~\cite{tsai2002code},~\cite{tauber2004verifiable},~\cite{tauber2004compiling} and used it to implement a number of protocols~\cite{georgiou2009automated}.
\cite{georgiou2009automated} lists six challenges when developing the compiler: ``Program structuring,'' ``IOA Programs and external services,'' ``Modeling procedure calls,'' ``Composing automata,'' ``Nondeterminism'', and ``Implementing datatypes.''
Programming written in IOA are compiled to Java and executed on hosts communicating using the MPI library.
IOA requires programs to presented in a ``node-channel'' form that allows them to be composed with built in ``mediators'' that implement external services such as MPI.
IOA lacks support for procedure calls which complicates modeling and interfacing with external services; especially when a procedure call blocks.
The IOA compiler composes automata statically using a ``composer'' resulting in a ``node automaton.''
The next action to be executed is determined by a schedule function that resolves the non-determinism in the schedule.

IOA and ioa++ represent two fundamentally different approaches to programming directly with I/O automata.
The development of a new language was necessary for IOA due to their focus on formal verification.
Ioa++ on the other hand uses the existing C++ language and compilers.
IOA encapsulates external services using built in ``mediators'' while ioa++ exposes operating system services via file descriptors relying on the programmer to ensure non-blocking operation.
IOA focuses on static composition to produce a single node program while ioa++ focuses on dynamic composition and allows concurrent execution.
Both IOA and ioa++ require the programmer to add code to schedule actions.

%% CORBA thread model

%% %% Lee~\cite{lee2006problem} also identifies a number of strategies for coping with thread-based programming including libraries, patterns, programming languages, and coordination languages.
%% %% In~\cite{schmitd2000patterns}, Schmidt et al. describe patterns for developing asynchronous and concurrent objects and systems.
%% %% The Split-C~\cite{culler1993parallel} and Cilk~\cite{blumofe1995cilk} add features to the C programming language for multi-threaded computation.
%% %% The Erlang~\cite{armstrong1996concurrent}

%% Ada

%% The

%% Events

%% Lynch

%% Garland

%% Ken Goldman

%% proactor/reactor
%% Observer