### 2012-02-23 papers/paper2/paper.tex

\documentclass{llncs}

\usepackage{cite}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{url}
\usepackage{amsmath}

\lstset{language=C++, basicstyle=\footnotesize}

\newif\ifjournal
\journalfalse

\begin{document}

\title{Implementing Reactive Systems Directly\\with I/O Automata in ioa++}
\author{Justin R. Wilson\inst{1} \and Christopher D. Gill\inst{1} \and Gruia-Catalin Roman\inst{2}}

\institute{
  Dept. of Computer Science and Engineering, 
  Washington University, St. Louis, MO
\and
  School of Engineering,
  University of New Mexico, Albuquerque, NM
  \{wilsonj,cdgill\}@cse.wustl.edu gcroman@unm.edu
}

\maketitle

\input{abstract}

\keywords{concurrency, I/O automata, components}

\input{introduction}

\input{component_model}

\input{design}

\input{programming_model}

\input{case_study}

\input{evaluation}

\input{related_work}

\input{conclusion}

\bibliographystyle{splncs03}
\bibliography{bibliography}{}

\end{document}